### 2022-06-22 omicron-generation.tex
homing (xiangshan)

Characterizing changes in the effective transmissibility over the course of infection---also known as the generation-interval distribution---is essential to comparing emerging SARS-CoV-2 variants.
Recent studies suggest that individuals infected with the Omicron variant develop symptoms earlier and transmit faster than those infected with the Delta variant.
However, these studies typically neglect the impact of population-level dynamics on their inference: 
when an epidemic is growing rapidly, we are more likely to find individuals who were infected recently, which biases us toward observing earlier transmission events.
When we account for this dynamical bias, we find that Omicron infections from the Netherlands at the end of December 2021 had similar incubation periods, but shorter generation intervals, compared to Delta infections from the same period.
Shorter generation intervals of the Omicron variant may be attributed to its higher effective reproduction number, which can cause faster local susceptible depletion around the contact network.

>>>>

Recent studies suggest that individuals infected with the Omicron variant develop symptoms earlier (shorter incubation period) and transmit faster (shorter generation interval) than those infected with the Delta variant.
However, these studies typically neglect population-level effects: 
when an epidemic is growing, a greater proportion of current cases were infected recently, biasing us toward observing faster transmission events.
Accounting for this dynamical bias, we find that Omicron infections from the Netherlands at the end of December 2021 had similar incubation periods, but shorter generation intervals, compared to Delta infections from the same period.
Shorter generation intervals of the Omicron variant might be due to its higher effective reproduction number, which can cause faster local susceptible depletion around the contact network.
